---
title: "CH 11: Assumptions - Part I"
output: pdf_document
---

\renewcommand{\vec}[1]{\mathbf{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 3, fig.width = 5, fig.align = 'center')
library(tidyverse) 
library(rstanarm)
library(arm)
library(gridExtra)
set.seed(10212020)
```

### Assumptions for Regression Models

The assumptions described in _Regression and Other Stories,_ are more broad than many textbooks. In order of importance,

\vfill

1. __Validity:__ 

\vfill

2. __Representativeness:__ 

\vfill

3. __Additivity and linearity:__ 

\vfill

4. __Independence of Errors:__ 

\vfill

5. __Equal Variance of Errors:__ 

\vfill

6. __Normality of Errors:__ 

\vfill
\newpage

What if the assumptions are violated??

\vfill


\vfill



\vfill



\vfill

### Plots of fitted model

For simple models with one continuous predictor and/or one categorical predictor, we have see how to fit the model with `geom_smooth`.

\vfill

With additional covariates in the model this becomes more challenging. Consider the candy dataset and a model

```{r}
candy <- read_csv("https://math.montana.edu/ahoegh/teaching/stat446/candy-data.csv") %>% 
  mutate(pricepercent = pricepercent - mean(pricepercent),
         sugarpercent = sugarpercent - mean(sugarpercent))
candy_model <- stan_glm(winpercent ~ pricepercent + sugarpercent, data = candy, refresh = 0)
print(candy_model)
plot(candy_model)
```

\vfill

- One option is to plot the response against each predictor holding the other continuous predictors constant and setting levels of categorical predictors. 

\vfill

```{r}
candy %>% 
  ggplot(aes(y = winpercent, x = pricepercent)) + 
  geom_point() + 
  geom_abline(intercept = candy_model$coefficients['(Intercept)'],
              slope = candy_model$coefficients['pricepercent'], 
              color = 'red') +
  labs(title = 'Model fit for winpercent vs. pricepercent \n for average sugarpercent') +
  theme_bw()
```

\newpage

### Residual Plots

Model fit can also be evaluated looking at residuals plots. 

\vfill

These plots should result in absence of patterns.

\vfill

#### Residual Plots from Fake Data

It is not always obvious (at least initially) what residual plots should look like and what variations could be expected when the model is indeed true.

\vfill

```{r, echo = F, fig.align = 'center', fig.height = 6, fig.width = 6}
n <- 80
x <- runif(n, -1, 1)
y <- rnorm(n, mean = x)
comb <- tibble(y = y, x = x)
model_fit <- comb %>% stan_glm(y~x, data = ., refresh = 0)
fig1 <- comb %>% ggplot(aes(y=y,x=x))+
  geom_smooth(formula = 'y~x', method = 'lm') + geom_point() + theme_bw() + labs(title = 'Y vs. X, simulated data')

fig2 <- tibble(resids = model_fit$residuals, fits = model_fit$fitted.values) %>% 
  ggplot(aes(y = resids, x = fits)) +
  geom_point() + theme_bw() +
  geom_smooth(formula = 'y~x', method = 'loess') + 
  labs(title = 'Residuals vs. fitted values')

grid.arrange(fig1, fig2)
```

\vfill

It can also be useful to create a panel of figures to explore residuals vs. each covariate.
\vfill

\newpage

